Scoring Data in R
===========================

## Looking at Grade Point Average, College Board Score and Highest Math Taken
## WSJ article:  "For Career Success, Study Those Quadratics"
## http://blogs.wsj.com/economics/2013/11/01/for-career-success-study-those-quadratics/

Student       | GPA    | College Board | Highest Math
:------------ | -----: | ------------: | :-----------
Ally Aardvark | 2.45   | 26            | Algebra
Billy Bison   | 3.82   | 30            | Calculus

### R code  (based on some code in R in Action)


```{r}
# install.packages("knitr")
# library(knitr)

Student <- c("Zsuzsa Polgar", "Vishy Anand", "Jose Capablanca", "Hou Yifan", "Magnus Carlsen")
GPA <- c(1.77, 2.45, 3.82, 2.75, 2.88)
BoardScore <- c(36, 26, 30, 28, 19)


candidates <- data.frame(Student, GPA, BoardScore,
    stringsAsFactors=FALSE)

z <- scale(candidates[,2:3])

score <- apply(z, 1, mean) 

candidates <- cbind(candidates, score)

y <- quantile(score, c(.8,.6,.4,.2))                                   

candidates$grade[score >= y[1]] <- "A"                                     
candidates$grade[score < y[1] & score >= y[2]] <- "B"
candidates$grade[score < y[2] & score >= y[3]] <- "C"
candidates$grade[score < y[3] & score >= y[4]] <- "D"
candidates$grade[score < y[4]] <- "F"

name <- strsplit((candidates$Student), " ")                                
lastname <- sapply(name, "[", 2)
firstname <- sapply(name, "[", 1)

candidates <- cbind(firstname,lastname, candidates[,-1])
candidates <- candidates[order(lastname,firstname),]

candidates

```

## Assignment 12

Your Task:  Build an application that uses some kind of scoring mechanism to normalize data

### Names
I wanted to create more data so I looked up a list of scientist names and copied those to names.csv

```{r}
# Read in a list of names and out of that create a fistname and lastname variable
names <- read.csv("..//Data//names.csv", stringsAsFactors=FALSE)
names <- names[,1]
names <- strsplit(names, " ")
firstname <- sapply(names, function(arg) arg[1])
lastname <- sapply(names, function(arg) arg[length(arg)])
```

### Create Data
```{r}
# Create random data for GPA, BoardScore, and Highest Math
set.seed(45)
GPA <- round(runif(length(names), min=0, max=4), digits=2)
BoardScore <- round(runif(length(names), min=1, max=36), digits=0)
MathLevels <- c("Algebra 1", "Algebra 2", "Geometry", "Trigonometry", "Calculus")
HighestMath <- MathLevels[round(runif(length(names), min=1, max=length(MathLevels)), digits=0)]
```

### Highest Math Score
```{r}
# Translate HighestMath into a MathCredit score to be used in ranking
# As a first pass I'm just going with a linear score
MathCredit <- sapply(HighestMath, function(arg) as.integer(which(MathLevels == arg)))
names(MathCredit) <- NULL
```

### Overall Score
```{r}
scores <- cbind(GPA, BoardScore, MathCredit)
z <- scale(scores[,1:3])
rank.score <- round(apply(z, 1, mean), digits=2)
rank <- rep(NA, length(rank.score))
rank[order(rank.score, decreasing=TRUE)] <- 1:length(rank.score)
```

### Output
```{r}
# Assemle this into a data frame
candidates <- data.frame(cbind(lastname, firstname, GPA, BoardScore, HighestMath, 
                               MathCredit, rank.score, rank))
row.names(candidates) <- NULL

options(width=120)
# Now display ordered by last name
candidates[order(lastname, firstname), ]

# We can also look at this by rank
candidates[order(rank), ]
```

Seeing that Einstein ranked 74 and Hawking ranked 174 this is probably not the most accurate system. ;-)

